Explorations to Optimize the Parareal Algorithm \\to Solve ODEs in Given Applications

### 2014-12-10 paper.tex
parareal picture

% Sets the page margins to 1 inch each side
\usepackage[margin=1in]{geometry}
\geometry{letterpaper}
%\addtolength{\oddsidemargin}{-0.875in} 
%\addtolength{\evensidemargin}{-0.875in} 
%\addtolength{\textwidth}{1.75in} 
%\addtolength{\topmargin}{-.875in} 
%\addtolength{\textheight}{1.75in}

@@

% Put your stuff here
%\twocolumn[
%\begin{@twocolumnfalse}
\maketitle
\begin{abstract}

We look to achieve strong scaling in parallelizing numerical ODE methods. We test both parallelism by result space as well as parallelism by time, with the Parareal algorithm, to solve various ODEs. We measure performance and compare in
terms of accuracy, speedup and efficiency. We provide some theoretical analysis for the Parareal algorithm and compare its tradeoff to paralleism by space, For the Parareal agorithm, we observe
light speedups in our small scale tests up to 64 processors on Harvard's
Odyssey cluster. We analyze possible reasons as to why our implementation
may not be ideal and some tradeoffs that are taken in the Parareal
implementation. Further optimizations are proposed and rationalized as next
steps including ports of to C++ using BLAS. Our code is written in python using mpi4py. We compare tests from various data sources from basic exponential ODEs to larger grids of ODEs to solve for in applications like diffusion and sound wave propagation.

@@

bksim TODO put a graph because maybe it'll help? the slides have a version

@@

The first t seconds comes from the first coarse approximation, without which the parareal algorithm degenerates to $g_{\Delta t}$. In order for there to be a speedup relative to the fine operator, we require
that $t + k(t + Qt/N) < Qt$, or $k < \frac{Q - 1}{1 + Q/N}$ or $N > \frac{Qk}{Q
- 1 - k}$. This can be tuned with by either adjusting the the quality factor or reducing the number of iterations, but the number of iterations, k, is important for error convergence to that of $g_{\textnormal{fine}}$. However, this optimal k is a problem-specific parameter and must be optimized for.

>>>>

% Sets the page margins to 1 inch each side
\usepackage[margin=1in]{geometry}
% force pictures to go where you want
\usepackage{float}

\geometry{letterpaper}
%\addtolength{\oddsidemargin}{-0.875in} 
%\addtolength{\evensidemargin}{-0.875in} 
%\addtolength{\textwidth}{1.75in} 
%\addtolength{\topmargin}{-.875in} 
%\addtolength{\textheight}{1.75in}

@@

% Put your stuff here
%\twocolumn[
%\begin{@twocolumnfalse}
\maketitle
\begin{abstract}
We look to achieve strong scaling in parallelizing numerical ODE methods. We test both parallelism by result space as well as parallelism by time, with the Parareal algorithm, to solve various ODEs.  We measure performance and compare in
terms of accuracy, speedup and efficiency. We provide some theoretical analysis for the Parareal algorithm and compare its tradeoff to paralleism by space, For the Parareal agorithm, we observe
light speedups in our small scale tests up to 64 processors on Harvard's
Odyssey cluster. We analyze possible reasons as to why our implementation
may not be ideal and some tradeoffs that are taken in the Parareal
implementation. Further optimizations are proposed and rationalized as next
steps including ports of to C++ using BLAS. Our code is written in python using mpi4py. We compare tests from various data sources from basic exponential ODEs to larger grids of ODEs to solve for in applications like diffusion and sound wave propagation.

@@

\begin{figure}[H]
\begin{center}
\includegraphics[width=0.75\linewidth]{parareal_visualization.png}
\caption{Graphical representation of the parareal scheme.}
\label{fig:pararealvis}
\end{center}
\end{figure}

@@

The first t seconds comes from the first coarse approximation, without which the parareal algorithm degenerates to $g_{\Delta t}$,  In order for there to be a speedup relative to the fine operator, we require
that $t + k(t + Qt/N) < Qt$, or $k < \frac{Q - 1}{1 + Q/N}$ or $N > \frac{Qk}{Q
- 1 - k}$. This can be tuned with by either adjusting the the quality factor or reducing the number of iterations, but the number of iterations, k, is important for error convergence to that of $g_{\textnormal{fine}}$.  However, this optimal k is a problem-specific parameter and must be optimized for.
